### Problem set Diamonds

***
```{r Loading data set}
library(ggplot2)
data(diamonds)
```

```{r Getting the basic information abot the data}
str(diamonds) 
# Number of observations ->53940
# Number of variables-> 10
# Number of ordered factors -> 3
# Letter for the best color of a diamond -> D
```

# Create a histogram of the price of
# all the diamonds in the diamond data set.
```{r Price histogram}
qplot(x = price, data = diamonds, binwidth = 30,
      color = I('black'), fill = I('#099DD9'))
```
# Price histogram summary
```{r Price histogram summary}
summary(diamonds$price)
```
I have come to a conclusion that most diamonds are not so expensive, the median is 2401, but based on the mean and median difference we can assume that there are also very expensive diamonds as well.

# Questions about the prices
```{r Prices}
nrow(subset(diamonds, price<500)) 
# Number of diamonds which cost less than 500$. -> 1729
nrow(subset(diamonds, price<250)) 
# Number of diamonds which cost less than 250$. -> 0
summary(diamonds$price >= 15000)
# Number of diamonds which cost more than 15000$. -> 1656
```
# Cheaper diamonds
# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won’t be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Submit your final code when you are ready.
```{r Largest peak}
qplot(x = price, data = diamonds, binwidth = 1,
      color = I('black'), fill = I('#099DD9')) + 
  scale_x_continuous(breaks = seq(600, 620, 1), limits = c (600, 620))
```

# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.

# TYPE YOUR CODE BELOW THE LINE
# ======================================================
```{r prices by cut}
qplot(x = price, data = diamonds) +
  facet_wrap(~cut)
```
# Which cut has the highest priced diamond?
```{r Prices by cut}
by(diamonds$price, diamonds$cut,max)
#Cut with the highest priced diamond. -> Premium
by(diamonds$price, diamonds$cut,min)
#Cut with the lowest priced diamond. -> Premium, Ideal
by(diamonds$price, diamonds$cut,median)
#Which cut has the lowest median price. -> Ideal
```

# In the two last exercises, we looked at
# the distribution for diamonds by cut.

# Run the code below in R Studio to generate
# the histogram as a reminder.

# ===============================================================
```{r Histogram}
qplot(x = price, data = diamonds) + facet_wrap(~cut)
```                                                                         
# ===============================================================

# In the last exercise, we looked at the summary statistics
# for diamond price by cut. If we look at the output table, the
# the median and quartiles are reasonably close to each other.

# diamonds$cut: Fair
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     337    2050    3282    4359    5206   18570 
# ------------------------------------------------------------------------ 
# diamonds$cut: Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     327    1145    3050    3929    5028   18790 
# ------------------------------------------------------------------------ 
# diamonds$cut: Very Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     336     912    2648    3982    5373   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Premium
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326    1046    3185    4584    6296   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Ideal
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326     878    1810    3458    4678   18810 

# This means the distributions should be somewhat similar,
# but the histograms we created don't show that.

# The 'Fair' and 'Good' diamonds appear to have 
# different distributions compared to the better
# cut diamonds. They seem somewhat uniform
# on the left with long tails on the right.

# Let's look in to this more.

# Look up the documentation for facet_wrap in R Studio.
# Then, scroll back up and add a parameter to facet_wrap so that
# the y-axis in the histograms is not fixed. You want the y-axis to
# be different for each histogram.

# If you want a hint, check out the Instructor Notes.

```{r Free scale y in facetwrap}
qplot(x = price, data = diamonds) + facet_wrap(~cut, scales = 'free_y')
```

# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

# Submit your final code when you are ready.

# ENTER YOUR CODE BELOW THIS LINE.
# ===========================================================================
```{r Histograms for price per carat}
qplot(x = price/carat, data = diamonds, binwidth = 0.01) +
  xlab("Price per carat") +
  ylab("Count of diamonds") +
  facet_wrap(~cut) +
  scale_x_log10()  

```

# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won’t be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# =================================================================
```{r Price of diamonds using boxplot}
qplot(x=color,y = price,data = diamonds,
      geom = "boxplot")+
  coord_cartesian(ylim = c(0,8000))
```
```{r Summary of prices based on color}
by(diamonds$price,diamonds$color, summary)
# First quartile for color D (25%) -> 911
# Third quartile for color D (75%) -> 4214
# First quartile for color J (25%) -> 1860
# Third quartile for color J (75%) -> 7695

by(diamonds$price,diamonds$color,IQR)
# IQR for diamonds with the best color -> 3303
# IRQ for diamonds with the worst color -> 5835
```

# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.

# Go to the discussions to
# share your thoughts and to discover
# what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.

# SUBMIT YOUR CODE BELOW THIS LINE
```{r Price per Carat based on color}
qplot(x=color,y = price/carat,data = diamonds,
  geom = "boxplot") + 
  coord_cartesian(ylim = c (1000, 5500))
```
```{r Carat size which has a count greater than 2000}
qplot(x = carat, data = diamonds, geom = 'freqpoly', xlab = 'Carat', binwidth = 0.01) +
  scale_y_continuous(breaks = seq(0,9000,1000))+
  scale_x_continuous(breaks = seq(0,3,0.1), limits = c(0, 3))
# Answer is -> Only 0.3 and 1.01 carat diamonds have a count greater than 2000.
```

### Gapminder data
```{r Load data}
titanic <- read.csv('titanic.csv', header = T, row.names = 1, check.names = F)

# Create box plot 
summary(titanic)

#Age of people
qplot(data = titanic, x = Age, binwidth = 1,
      color = I('black'), fill = I('#099DD9')) + xlab("Age of person") + ylab("Number of people") + scale_x_continuous(breaks = seq(0, 80, 5))

#Price of ticket based on sex
qplot(data = titanic, x = Sex, y = Fare, geom = "boxplot", 
      color = I('black'), fill = I('orange')) + coord_cartesian(ylim = c(0, 150))

#Placement in the class based on sex
qplot(data = titanic, x = Sex, y = Pclass, geom = "boxplot", 
      color = I('black'), fill = I('orange'))

#Price graph
qplot(x = Fare, data = titanic, geom = 'freqpoly') +
  scale_x_continuous(breaks = seq(0,500,50)) +
  scale_y_continuous(breaks = seq(0,320,50))
```
### Birthday data
```{r Birthday data analysis}
birthdays <- read.csv("birthdaysExample.csv", header = T, check.names = F)
install.packages("glue")
install.packages("stringi")
install.packages("stringr")
install.packages("lubridate")
library("lubridate")

birthdays$dates <- mdy(birthdays$dates)
birthdays$month <- month(birthdays$dates)
birthdays$day <- day(birthdays$dates)

# How many people have the same birthday as you?
# (Reserve time with them or save money to buy them a gift!)
mybirthday <- subset(birthdays, month==01 & day == 18)
nrow(mybirthday)
Answer: 1
# Which month contains the most number of birthdays?
qplot(x = month, data = birthdays, binwidth = 1,
      color = I('black'), fill = I('#099DD9')) + 
  scale_x_continuous(breaks = seq(1, 12, 1))
Answer: 3
# How many birthdays are in each month?
table(birthdays$month)

# Which day of the year has the most number of birthdays?
qplot(x = day, data = birthdays, binwidth = 1,
      color = I('black'), fill = I('#099DD9')) + 
  scale_x_continuous(breaks = seq(1, 31, 1))
Answer: 14

# Do you have at least 365 friends that have birthdays on everyday
# of the year?
length(unique(birthdays$dates))
Answer: No
```
