---
output: html_document
editor_options: 
  chunk_output_type: console
---
Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes: People guessed that their audiance size is smaller than it actually it is and also people guessed round numbers like 50, 100 and even 200.

***

### Scatterplots
Notes: 

```{r Scatterplots}
library(ggplot2)
facebook <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = facebook)
qplot(age, friend_count, data = facebook)
```

***

#### What are some things that you notice right away?
Response: The people who are younger also have more friends, then there are exteremes like very old people (age > 90) having a bunch of friends.

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = facebook) + geom_point() + xlim(13,90)

summary(facebook$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = facebook) + 
  geom_jitter(alpha = 1/20) + 
  xlim(13,90) 
```

#### What do you notice in the plot?
Response: Young users have a lot of friends, and a lot of young people have friend count less than 1000.

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = facebook) + 
  geom_jitter(alpha = 1/20) + 
  xlim(13,90)
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = facebook) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13,90) +
  coord_trans(y = 'sqrt')
```

#### What do you notice?
The graph is more readable now.
***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = facebook) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13,90) +
  coord_trans(y = 'sqrt')
```
A lot of people initated a lot of friendships when they were young. Also we can see the line going up at about age of 69, these are probably the fake accounts.
***

### Overplotting and Domain Knowledge
Notes: The lecturer said, that if we transformed the number to a percentage it is more clear how many people really saw the post so compared to percieved audience, it is clear that peolple are underestimating again.

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr')
library('dplyr')
suppressMessages(library(dplyr))

facebook.fc_by_age <- facebook %>% 
  group_by(age) %>%
  summarise(friend_count_mean=mean(friend_count),
            friend_count_median=median(friend_count),
            n=n()) %>%
  arrange(age)
head(facebook.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x=age, y=friend_count_mean), data=facebook.fc_by_age) +
  geom_line() +
  ggtitle("Average Friend Count vs. Age") +
  labs(x="Age", y="Average Friend Count") +
  scale_x_continuous(breaks=seq(10, 115, 5)) +
  scale_y_continuous(breaks=seq(50, 500, 50))
```

***

### Overlaying Summaries with Raw Data
Notes: 

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = facebook) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') + 
  xlim(13,90) + 
  coord_trans(y = 'sqrt') +
  stat_summary(geom = "line", fun = "mean") +
  stat_summary(geom = "line", fun = "quantile", fun.args = list(probs = .1), linetype = 2, color = 'blue') +
  stat_summary(geom = "line", fun = "quantile", fun.args = list(probs = .5), linetype = 2, color = 'blue') +
  stat_summary(geom = "line", fun = "quantile", fun.args = list(probs = .9), linetype = 2, color = 'blue') +
  coord_cartesian(xlim = c(13, 90), ylim = c(0, 1000))

```

#### What are some of your observations of the plot?
Response: Most people of young age have a lot of friends, but then it starts to decrease. Also we see a big spike for people aged 69, these people are not actually 69 years old. Then we can also see an increase in the friends size to the older age, which may also implicate that these data are fake.

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes: We can see in the darker blue color the histogram for people who overestimated their audience size and in the lighter blue the people who have overestimated the audience size. Again we can see that people underestimate a lot. In the data for asking people in the longer period of time so (how many people saw your post in last month). The predictions get closer to reality, but not too much.

***

### Correlation
Notes:

```{r Correlation}
cor.test(facebook$age, facebook$friend_count, method = 'pearson')
with(facebook, cor.test(age, friend_count, method = 'pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response: -0.027

***

### Correlation on Subsets
Notes: Based on previous calculation we know that relationship between age and friend count is not linear.

```{r Correlation on Subsets}
with(subset(facebook, age <= 70), cor.test(age, friend_count))
```
This gives us a different summary statistic, r = -0.1717245. This tells a different story about a negative relationship between age and friend count. As age increases, we can see that friend count decreases.
***

### Correlation Methods
Notes: Single number coefficients like correlations are useful, but they are not substitute for a scatter plot and computing conditional summaries.

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x=www_likes_received, y=likes_received), data=facebook) +
  geom_point(alpha = .1)

```

***

### Strong Correlations
Notes: 

```{r Strong Correlations}
ggplot(aes(x=www_likes_received, y=likes_received), data=facebook) +
  geom_point() +
  xlim(0, quantile(facebook$www_likes_received, 0.95)) + 
  ylim(0, quantile(facebook$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(subset(facebook), cor.test(www_likes_received, likes_received))
```

Response: 0.948

***

### Moira on Correlation
Notes:
A lot of data is correlated with each other, in Moiras case these data show engagement on Facebook
assumption of regression is that these variables are independent from each other, because we cannot define straight away which one is driving the phenomenom of correlation.
***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages("alr3_2.0.8.tar.gz", repos = NULL, type="source")

install.packages('car')
library(alr3)
data(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(Month, Temp),data=Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

I think it is 0.08

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

0.057

```{r Noisy Scatterplots}
with(data = Mitchell, cor.test(Month, Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x=(Month%%12), Temp),data=Mitchell) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response: That in summer the temperature gets higher.

Watch the solution video and check out the Instructor Notes!
Notes: We need to also size the graph appropriately to see the story behind it.

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
facebook$age_with_months <- facebook$age_with_months <- facebook$age + (1 - facebook$dob_month / 12)
```

***

### Age with Months Means

```{r Age with Months Means}
facebook.fc_by_age_months <- facebook %>% 
  group_by(age_with_months) %>%
  summarise(friend_count_mean=mean(friend_count),
            friend_count_median=median(friend_count),
            n=n()) %>%
  arrange(age_with_months)
head(facebook.fc_by_age_months)
```

Programming Assignment
```{r Programming Assignment}
ggplot(aes(x=age_with_months, y=friend_count_mean), data=subset(facebook.fc_by_age_months, age_with_months < 71)) +
  geom_line() +
  ggtitle("Average Friend Count vs. Age with months") +
  labs(x="Age with months", y="Average Friend Count") 
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

p1 <- ggplot(aes(x=age, y=friend_count_mean), data = subset(facebook.fc_by_age, age < 71)) +
  geom_line() 


p2 <- ggplot(aes(x=age_with_months, y=friend_count_mean), data=subset(facebook.fc_by_age_months, age_with_months < 71)) +
  geom_line() 

p3 <- ggplot(aes(x = round(age/5)*5, y = friend_count),
             data = subset(facebook, age < 71)) +
  geom_line(stat="summary", fun = "mean")

library(gridExtra)
grid.arrange(p2,p1,p3, ncol = 1)
  
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x=age, y=friend_count_mean), data = subset(facebook.fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()


p2 <- ggplot(aes(x=age_with_months, y=friend_count_mean), data=subset(facebook.fc_by_age_months, age_with_months < 71)) +
  geom_line() +
  geom_smooth()

p3 <- ggplot(aes(x = round(age/5)*5, y = friend_count),
             data = subset(facebook, age < 71)) +
  geom_line(stat="summary", fun = "mean")
grid.arrange(p2,p1,p3, ncol=1)

library(gridExtra)
grid.arrange(p2,p1,p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes: We dont have to choose "correct plot", because we can get more interesting data from the different graphs. Different types of graphs are useful. When we present our work we may want to choose some graph, which represents our conclusion the best.

***

### Analyzing Two Variables
Reflection: This lesson gave me understanding of how tricky can correlation be. I have learned how to visualize these correlations in scatter plots and different types of plots and also to find some interesting things in plots, based on different visualization. I have also learned that there are different models for data, and how we can choose to trade of something to better understand other thing as in the last example of data smothering. I also learned how we can make bad assumptions based on graphs and how to deal with it, or check it the other way so i dont make mistake in my assumptions.

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

