---
output: html_document
editor_options: 
  chunk_output_type: console
---

### Problem set Diamonds

------------------------------------------------------------------------

```{r Loading data set}
library(ggplot2)
data(diamonds)
```

```{r Scatterplot of price vs x}
ggplot(aes(x = x, y = price), data = diamonds) + geom_point()
```

### What are your observations about price vs x?

The higher is the x the higher is also the price, but it goes up exponentialy. We can also se some weird values on the x scale in the number 0. Maybe there are just bad values or some extereme values. 

```{r Correlation}
with(diamonds, cor.test(price, x, method = 'pearson'))
```
### What is the correlation between price and x?
0.88
```{r Correlation}
with(diamonds, cor.test(price, y, method = 'pearson'))
```
### What is the correlation between price and y?
0.87
```{r Correlation}
with(diamonds, cor.test(price, z, method = 'pearson'))
```
### What is the correlation between price and z?
0.88

### Scatterplot price vs depth
```{r Scatterplot of price vs depth}
ggplot(data = diamonds, aes(x = depth, y = price)) + 
  geom_point()
```

### Modified plot
```{r Modified plot}
ggplot(data = diamonds, aes(x = depth, y = price)) + 
  geom_point(alpha=0.01) +
  scale_x_continuous(breaks=seq(42,80,2))
```

### Based on the scatterplot of depth vs price, most diamonds are between what values?
Between %8 and 64.

### What is the correlation between depth and price?
0.87
```{r Correlation}
with(diamonds, cor.test(depth, price, method = 'pearson'))
```

### Create a scatterplot of price vs carat and omit the top 1% of price and carat values.
```{r Scatterplot price vs carat}
ggplot(data = diamonds, aes(x = carat, y = price)) + 
  xlim(0, quantile(diamonds$carat, 0.99)) + 
  ylim(0, quantile(diamonds$price, 0.99)) +
  geom_point()
```

### Create a new variable -> volume.
```{r Create volume variable}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z
```

### Create a scatterplot of price vs. volume (x * y * z).
```{r Scatterplot price vs volume}
ggplot(data = diamonds, aes(x = volume, price)) +
  geom_point()
```
### Observations based on this graph:
The price goes exponentially up based on a volume of diamond.

```{r Correlation of price vs volume, with omited values}
with(subset(diamonds, diamonds$volume > 0 & diamonds$volume <= 800), cor.test(price, volume))
```

```{r Adjusted scatterplot price vs volume}
ggplot(data = subset(diamonds, diamonds$volume > 0 & diamonds$volume <= 800), aes(x = volume, price)) +
  geom_point(alpha = 0.01) + 
  geom_smooth(method = 'lm', color = 'blue')
```

### New data frame diamonds ByClarity
```{r New data frame}
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
diamondsByClarity <- group_by(diamonds, clarity) %>%
  summarise(mean_price=mean(price),
            median_price=median(price),
            min_price=min(price),
            max_price=max(price),
            n=n())
```

### Bar charts of Mean price
```{r Bar charts of mean price}
library(dplyr)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(diamonds_mp_by_clarity, aes(x = clarity, y = mean_price)) + geom_bar(stat = 'identity')
p2 <- ggplot(diamonds_mp_by_color, aes(x = color, y = mean_price)) + geom_bar(stat = 'identity')

library(gridExtra)
grid.arrange(p1, p2, ncol = 1)
```
### Results of observation
I see that based on the clarity price is not differentiating a lot. However based on a colort we can see that closer to the color J the price gets higher linearly.

### Gapminder data
``` {r Consumption of sugar per person}
install.packages('xlsx')
install.packages('reshape2')
library('xlsx', quietly=TRUE)
library(reshape2)
sugar <- read.xlsx('sugar.xlsx', sheetIndex=1)
sugar[ncol(sugar)] <- NULL
sugar <- sugar[!is.na(sugar[[1]]),]
colnames(sugar)[1] <- 'country'
sugar$country <- gsub(' ','_',sugar$country)
countries <- as.character(sugar$country)
years <- seq(1961,2017)
sugar2 <- as.data.frame(t(sugar[,-1]))
colnames(sugar2) <- countries
sugar2['year'] <- years
ggplot(sugar2, aes(x=year, y=USA)) + geom_line() + xlab('Year') + ylab('Sugar per Person (g per day)') + ggtitle('Sugar Consumption in the United States')
```

```{r Sugar per person vs. year for all countries}
sugar3 <- melt(sugar2, id.vars='year', na.rm=TRUE, variable.name='country')
colnames(sugar3)[3] <- 'amount'
ggplot(subset(sugar3, sugar3$amount < 500), aes(x=year, y=amount)) + 
  geom_point(alpha=1/3, position = position_jitter(h=0), color='orange') + 
  stat_summary(geom ='line', fun = 'mean', color = 'red') +
  stat_summary(geom ='line', fun = 'quantile', fun.args=list(probs=0.1), linetype=2, color='brown') + 
  stat_summary(geom ='line', fun = 'quantile', fun.args=list(probs=0.9), linetype=2, color='blue') +
  stat_summary(geom ='line', fun = 'median', color='blue') + 
  xlab('Year') + 
  ylab('Sugar per Person (g per day)') +
  ggtitle('Sugar Consumption Worldwide from 1961-2017')

```