---
output: html_document
editor_options: 
  chunk_output_type: console
---
Lesson 5
========================================================

### Multivariate Data
Notes: In this lesson we will examine the realtionships between multiple variables. 

***

### Moira Perceived Audience Size Colored by Age
Notes: Moira made an addition of color to the age variable to maybe examine the dependency that with age people become more accurate, but as she said she failed.

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library('dplyr')
suppressMessages(library(dplyr))
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count=mean(friend_count),
            median_friend_count=median(friend_count),
            n=n()) %>%
  ungroup()
head(pf.fc_by_age_gender)


ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot()
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(pf.fc_by_age_gender, aes(x=age,y=median_friend_count, color=gender)) + geom_line()
```

***

### Thinking in Ratios
Notes: So we need to count the difference in ratios better than in specific numbers.

***

### Wide and Long Format
Notes: It is useful to group data more or less based on our investigation.

***

### Reshaping Data
Notes:

```{r}
install.packages("tidyr")
library(tidyr)

spread(subset(pf.fc_by_age_gender, 
       select = c('gender', 'age', 'median_friend_count')), 
       gender, median_friend_count)
library(reshape2)
pf.fc_by_age_gender_wide <- dcast(pf.fc_by_age_gender, age~gender, value.var='median_friend_count')
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(pf.fc_by_age_gender_wide, aes(age, y=female/male)) + geom_line() +
  geom_hline(yintercept=1, alpha=0.3, linetype=2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- 2014-ceiling(pf$tenure/365)

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004,2009,2011,2012,2014), right=TRUE)
table(pf$year_joined.bucket, useNA='ifany')

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun = 'median')
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun = 'median') +
  geom_line(stat='summary', fun = 'mean', linetype=2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
summary( with(subset(pf, tenure>0), friend_count/tenure) )
```

***

### Friendships Initiated
Notes:

What is the median friend rate?
0.2205
What is the maximum friend rate?
417
```{r Friendships Initiated}
ggplot(subset(pf, tenure>0), aes(x=tenure, y=friendships_initiated/tenure, color=year_joined.bucket)) + 
  geom_line(stat='summary', fun = 'mean')
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

library(gridExtra)
p1 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p4 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

grid.arrange(p1,p2,p3,p4, ncol=1)
```

***

### Sean's NFL Fan Sentiment Study
Notes: He is a NFL Fan himself and he feels the ups and lows as all the fans do so he decided maybe we can visualize this somehow. When we apply some trade of of time and data measurements we lose differencies in time but get overall more overview.

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

yo$id <- factor(yo$id)
str(yo)

qplot(data = yo, x = price)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
yo <- transform(yo, all.purchases=strawberry+blueberry+pina.colada+plain+mixed.berry)

```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(data = yo, aes(x = time, y = price)) + geom_point(alpha = 0.1, position=position_jitter())
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(1)
sample.ids <- sample(levels(yo$id), 16)

ggplot(subset(yo, id %in% sample.ids), aes(x=time, y=price)) +
  facet_wrap(~id) + 
  geom_line() + 
  geom_point(aes(size=all.purchases), pch=1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r Scatter plot matrix}
install.packages('GGally')
library(GGally)
theme_set(theme_minimal(20))

# set the seed for reproducible results
set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```
***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection: This lesson presented the concept of adding more variables to the summaries in the previous lessons. I learned how to make comparison of many variables and was introduced to the principle of heat maps. I learned how to reshape (group) data and learned how to make wide and long data.

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

